Distrivia: A Distributed Trivia Game

### 2011-04-16 papers/phase-two/paper.tex
Further updates to phase-two paper

\section {What We've Done}
Much of our system has been implemented and is in the debugging phase. The server apps and the web client are nearly complete with some revisions happening as necessary. The extent of what has been finished for each part of our system is explained below.
\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cloud \cite{aec}. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly. This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.
\subsection {Traffic}
We have two load sharers set up. Both sharers are currently configured to
divide the requests among the three active servers and, if need be, the two hot spares. The duplicate sharers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load sharers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load sharers are setup to attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.
Currently we are using the NGINX \cite{nginx} in a reverse proxy configuration
for our load sharing needs.

@@

\subsection {Messages}
We are currently using the https protocol for our messages. This provides a secure
and usable messaging system that we have designed our client and server APIs to use.
The actual messages are sent and received in two way's. Initially a user session key is
generated using the UUID \cite{uuid} algorithm and returned to the user on successful
authentication. This session key is then passed back to the server with every in coming
message. Messages are sent to the server using POST and GET request methods supported
by the HTTP protocol. Responses from the server are in the form of JSON \cite{json} objects.
JSON is a small, easily parseable format, representing objects in a human readable format as
they would be represented in JavaScript. With the web client the JSON object returned from
the server is instantly usable, however for the android client we have to parse the JSON object
into a native Java object. The Android SDK provides libraries to do this, located in the
org.json \cite{orgjson} package name space.
\section {What's Left}
\subsection {Server}
The server is currently almost feature complete, but still needs a lot of
testing. The only features it does not currently support is any sort of notion
of a private game. Client registration, and login are all supported. A client can
join a game, retrieve questions for that game, and answer questions. Both a
global leader board and individual per game leader board can also be requested.
As a group we have already found several inconsistencies and things we would
like to change with how server API requests are structured. We have also thought of
way's we can combine multiple requests into one, reducing the possible latency
and overall communication time of multiple requests. To fix these
inconsistencies members of the group worked to create a formal specification
for the request protocol of each possible request. This allows the server and client
programmers to work towards a common goal with out either making inappropriate
assumptions about how the system will work. We are currently in the middle
of refactoring the server to use the new requests detailed in the API.

\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
While private games are implemented on the server, 
Also, when a user clicks on a button to do something like log in, the button should be disabled while waiting for a response,
alternatively it needs to be able to handle sending out two calls for the same request and not getting confused if it gets two back.
\subsection {Android}
EVERYTHING
So far, the only connection to the server that the android client has is when it
registers/logs in. We still need to connect the rest of the android client to
the server API to make sure all the communication between the android client
and the server are working. This includes joining public/private rounds,
competing in these rounds, and viewing leader boards. While the UI is nearly
finished on the android client, nearly all of the connections to the servers
has yet to be done. Additionally, the android client needs to have more user
feedback and increased ability for the users to control which activity they
are in.
\section {Meeting the Requirements}

>>>>

\section {What We've Done}
Much of our system has been implemented and is in the debugging phase. The server apps and the web client are nearly complete with some revisions happening as necessary. The extent of what has been finished for each part of our system is explained below.
\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cloud \cite{aec}. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly. This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.

\subsection {Server Applications}
The server is currently almost feature complete, but still needs a lot of
testing. Client registration, and login are all supported. A client can
join a game, retrieve questions for that game, and answer questions. Both a
global leader board and individual per game leader board can also be requested. We recently refactored our API to be more universal. The servers and clients now work on a more consistent basis where the clients all expect the same information and use the same calls.

\subsection {Traffic}
We have two load sharers set up. Both sharers are currently configured to
divide the requests among the three active servers and, if need be, the two hot spares. The duplicate sharers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load sharers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load sharers are setup to attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.
Currently we are using the NGINX \cite{nginx} in a reverse proxy configuration
for our load sharing needs.

@@

\subsection {Messages}
We are currently using the https protocol for our messages. This provides a secure
and usable messaging system that we have designed our client and server APIs to use.
The actual messages are sent and received in two way's. Initially a user session key is
generated using the UUID \cite{uuid} algorithm and returned to the user on successful
authentication. This session key is then passed back to the server with every in coming
message. Messages are sent to the server using POST and GET request methods supported
by the HTTP protocol. Responses from the server are in the form of JSON \cite{json} objects.
JSON is a small, easily parseable format, representing objects in a human readable format as
they would be represented in JavaScript. With the web client the JSON object returned from
the server is instantly usable, however for the android client we have to parse the JSON object
into a native Java object. The Android SDK provides libraries to do this, located in the
org.json \cite{orgjson} package name space.

\section {What's Left}
\subsection {Server}
The server is nearly completed. The server application just needs to have some tweaks and further testing.
\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
While private games are implemented on the server, 
Also, when a user clicks on a button to do something like log in, the button should be disabled while waiting for a response,
alternatively it needs to be able to handle sending out two calls for the same request and not getting confused if it gets two back.
\subsection {Android}
We've started to thread the network calls to correctly register UI updates to the application. This is completed for most of the activities. There are really two big things left to implement on the android client: private games and competing in a round. Competing in a round is nearly finished, we just need to refresh the questions/answers and post the users answers to the server. Besides these two things, we just need to solidify the user interface, make it work with the back button with popping activities off the stack and other minor updates.

\section {Meeting the Requirements}
